The large scale clustering of fast radio bursts in dispersion measure space

### 2015-03-21 writeup/dm_space_distortions.tex
started code to do integrals.

For dark matter, ignoring time dependance (redshift evolution)
of the power spectrum:
\begin{align}
\delta_{\ell\ell'}\delta_{mm'}C_\ell(\chi, \chi')
    &= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \langle \delta(\hat n \chi) \delta(\hat n' \chi') \rangle
        \\
    &= gg'aa'
        \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \int\frac{\ud^3\vec k}{(2 \pi)^3} \frac{\ud^3\vec k'}{(2 \pi)^3}
        e^{i\vec k \cdot \vec x} e^{i\vec k' \cdot \vec x'}
        \langle \delta(\vec k, 0) \delta(\vec k', 0) \rangle
        \\
    &= gg'aa'
        \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \int\frac{\ud^3\vec k}{(2 \pi)^3} 
        e^{i\vec k \cdot (\vec x - \vec x')} P(k, 0)
        \\
    &= gg'aa'
        \int\frac{\ud^3\vec k}{(2 \pi)^3} P(k, 0)
        \int\ud\Omega Y_{\ell m}(\hat n)
        e^{i\vec k \cdot \vec x}
        \int\ud\Omega' Y_{\ell' m'}(\hat n')
        e^{-i\vec k \cdot \vec x'}
        \\
    &= gg'aa'
        \int\frac{\ud^3\vec k}{(2 \pi)^3} P(k, 0)
        \left[(4 \pi) i^\ell j_\ell(k\chi) Y_{\ell m}^*(\hat k)\right]
        \left[(4 \pi) i^{-\ell'} j_{\ell'}(k\chi') Y_{\ell' m'}(\hat
        k)\right]
        \\
    &= gg'aa'\frac{2}{\pi}
        \int\ud k k^2 j_\ell(k\chi) j_{\ell'}(k\chi')P(k, 0)
        \int\ud \Omega_k i^{(\ell - \ell')}
        Y_{\ell' m'}^*(\hat k) Y_{\ell m}(\hat k)
        \\
    &= \delta_{\ell\ell'}\delta_{mm'}gg'aa'\frac{2}{\pi}
        \int\ud k k^2  j_\ell(k\chi) j_{\ell}(k\chi')P(k, 0)
        \\
C_\ell(\chi,\chi') 
    &= gg'aa'\frac{2}{\pi}
        \int\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P(k, 0)
\end{align}

@@

Define $\delta_d$ as:
\begin{equation}
    \delta_d(\hat n \chi) = \int_0^\chi \ud \chi' \delta_e(\hat n \chi')
\end{equation}
which doesn't have the same units as over-density, but that is okay.
\begin{equation}
C^{dd}_\ell(\chi,\chi') 
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''g''a''
    \int_0^{\chi'}\ud\chi'''
    g'''a'''
    \int\ud k k^2 j_\ell(k\chi'') j_{\ell}(k\chi''')P_{ee}(k, 0)
\end{equation}
At this point we could again to the transformation to $C_{\ell\chi\chi'}$
leaving the expression as a triple integral. This is not as computationally
costly as might be expected at first glance, since the $\chi''$ and $\chi'''$
integrals are almost the same for all $\chi$, $\chi'$ pairs with only the
limits of integration changing. Thus the full set of $C_{\ell\chi\chi'}$'s can
be computed with a cumulative sum, only performing one new integral (over $k$)
for each element.

@@

Finally, the cross terms:
\begin{equation}
C^{dx}_\ell(\chi,\chi')
    =
    g'a'
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''g''a''
    \int\ud k k^2 j_\ell(k\chi'') j_{\ell}(k\chi')P_{ex}(k, 0)
\end{equation}
Or under Limber:
\begin{align}
C^{dx}_\ell(\chi,\chi')
    &=
    g'a'
    \int_0^\chi\ud\chi''g''a''
    \frac{1}{\bar\chi^2}
    \delta(\chi'' - \chi')
    P_{ex}(\ell/\bar\chi, 0)
    \\
    &=
    \left(\frac{g'a'}{\chi'}\right)^2
    P_{ex}(\ell/\chi', 0)
    \qquad \textrm{if $\chi' < \chi$ and 0 otherwise.}
    \\
C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
    &=
    \left(\frac{g^ma^m}{\chi^m}\right)^2
    P_{ex}(\ell/\chi^m, 0)
\end{align}

>>>>

For any statistically stationary tracers $x$ and $y$,
ignoring time dependance (redshift evolution)
of the power spectrum:
\begin{align}
    \delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    &= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \langle \delta^x(\hat n \chi) \delta^y(\hat n' \chi') \rangle
        \\
    &=
        \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \int\frac{\ud^3\vec k}{(2 \pi)^3} \frac{\ud^3\vec k'}{(2 \pi)^3}
        e^{i\vec k \cdot \vec x} e^{i\vec k' \cdot \vec x'}
        \langle \delta(\vec k) \delta(\vec k') \rangle
        \\
    &=
        \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \int\frac{\ud^3\vec k}{(2 \pi)^3} 
        e^{i\vec k \cdot (\vec x - \vec x')} P_{xy}(k)
        \\
    &=
        \int\frac{\ud^3\vec k}{(2 \pi)^3} P_{xy}(k)
        \int\ud\Omega Y_{\ell m}(\hat n)
        e^{i\vec k \cdot \vec x}
        \int\ud\Omega' Y_{\ell' m'}(\hat n')
        e^{-i\vec k \cdot \vec x'}
        \\
    &=
        \int\frac{\ud^3\vec k}{(2 \pi)^3} P_{xy}(k)
        \left[(4 \pi) i^\ell j_\ell(k\chi) Y_{\ell m}^*(\hat k)\right]
        \left[(4 \pi) i^{-\ell'} j_{\ell'}(k\chi') Y_{\ell' m'}(\hat
        k)\right]
        \\
    &=
        \frac{2}{\pi}
        \int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell'}(k\chi')P_{xy}(k)
        \int\ud \Omega_k i^{(\ell - \ell')}
        Y_{\ell' m'}^*(\hat k) Y_{\ell m}(\hat k)
        \\
    &= \delta_{\ell\ell'}\delta_{mm'}\frac{2}{\pi}
        \int_0^\infty\ud k k^2  j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k)
        \\
    C^{xy}_\ell(\chi,\chi') 
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k)
\end{align}

It is convienient to reparameterize this in terms of
$\bar\chi \equiv (\chi + \chi') /2$ and $\Delta \chi \equiv \chi' - \chi$.
We expect the power spectrum to vary slowly as a function of $\bar\chi$ and
very quickly as a function of $\Delta\chi$. However it will be highly
suppressed for $\ell \gg \bar\chi/\Delta\chi$. These facts will be very
helpfull in subsequent numerical integrals.

Also since smallish value of $\Delta\chi$ are of interest, this leads us to
an obviouse way to include redshift dependance of the
power spectrum:
\begin{equation}
C^{xy}_\ell(\bar\chi,\Delta\chi) 
    = \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,\bar\chi) \qquad \Delta\chi \ll 1/aH
\end{equation}

Define $\delta_d$ as:
\begin{equation}
    \delta_d(\hat n \chi) = \int_0^\chi \ud \chi' \delta_e(\hat n \chi')
\end{equation}
which doesn't have the same units as over-density, but that is okay.
\begin{equation}
C^{dd}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''
    \int_0^{\chi'}\ud\chi'''
    \int_0^\infty\ud k k^2 j_\ell(k\chi'') j_{\ell}(k\chi'''
    )P_{ee}(k, (\chi''+\chi''')/2)
\end{equation}

For numerical integration, perform the following change of variables to make
things much easier.
\begin{equation}
C^{dd}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^{(\chi' + \chi)/2}\ud\bar\chi'
    \int_{\max(-2\bar\chi', -2(\chi -\bar\chi'))}^{\min(2\bar\chi', 2(\chi' -
    \bar\chi'))}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\bar\chi' - \Delta'/2))
    j_{\ell}(k(\bar\chi' + \Delta'/2))
    P_{ee}(k, \bar\chi')
\end{equation}
At this point we could again to the transformation to $C_{\ell\chi\chi'}$
leaving the expression as a triple integral. This is not as computationally
costly as might be expected at first glance, since the $\chi''$ and $\chi'''$
integrals are almost the same for all $\chi$, $\chi'$ pairs with only the
limits of integration changing. Thus the full set of $C_{\ell\chi\chi'}$'s can
be computed with a cumulative sum, only performing one new integral (over $k$)
for each element.

Finally, the cross terms:
\begin{equation}
C^{dx}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''
    \int_0^\infty\ud k k^2 
    j_\ell(k\chi') j_{\ell}(k\chi'')P_{ex}(k, (\chi' + \chi'')/2)
\end{equation}
\begin{equation}
    C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''
    \int_0^\infty\ud k k^2 
    j_\ell(k\chi') j_{\ell}(k\chi'')P_{ex}(k, (\chi' + \chi'')/2)
\end{equation}

with the following change of variables for numerical integration:
\begin{align}
C^{dx}_\ell(\chi,\chi')
    =&
    \frac{2}{\pi}
    \int_{-\chi'}^{\chi - \chi'}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k\chi')
    j_{\ell}(k(\chi'+\Delta'))
    P_{ex}(k, \chi' + \Delta'/2)
    \\
C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
    \approx&
    \frac{2}{\pi}
    \int_{-\chi'}^{\chi - \chi'}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi'-\Delta/2))
    j_{\ell}(k(\chi'+\Delta'/2))
    P_{ex}(k, \chi')
    \nonumber\\
    &+
    \frac{2}{\pi}
    \int_{-\chi}^{\chi' - \chi}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi-\Delta/2))
    j_{\ell}(k(\chi+\Delta'/2))
    P_{ex}(k, \chi)
\end{align}
The second line exploits the fact that for $\ell \gg 1$, $\chi \gg \chi'/\ell$ and
$\chi' \gg \chi/\ell$, only $\Delta'\ll\chi$ contributes significantly to the
integral. Also the expressions are very weak functions of $\chi$ and thus
may be shifted by $\Delta/2$ without incurring much error.

@@

@@

Or under Limber:
\begin{align}
C^{dx}_\ell(\chi,\chi')
    &=
    g'a'
    \int_0^\chi\ud\chi''g''a''
    \frac{1}{\bar\chi^2}
    \delta(\chi'' - \chi')
    P_{ex}(\ell/\bar\chi, 0)
    \\
    &=
    \left(\frac{g'a'}{\chi'}\right)^2
    P_{ex}(\ell/\chi', 0)
    \qquad \textrm{if $\chi' < \chi$ and 0 otherwise.}
    \\
C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
    &=
    \left(\frac{g^ma^m}{\chi^m}\right)^2
    P_{ex}(\ell/\chi^m, 0)
\end{align}
